---
title: "EVALUATION OF RECOMMENDER SYSTEMS"
author: "Alain Kuiete Tchoupou"
date: "April 9, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the libraries

```{r}
library(recommenderlab)  # Matrix/recommender functions
library(dplyr)  # Data manipulation
library(tidyr)  # Data manipulation
library(ggplot2)  # Plotting
library(tictoc)  # Operation timing
```
```{r}
set.seed(42)
```

## Subsetting Dataset
Now let’s see the process of reducing the data from the main dataset

Import original file and select sample for project

```{r}
ratings <- read.csv('C:\\DATA612\\DATA612ASSINGMENTS-master\\ml-20m\\ratings.csv')
```

```{r}
ratings <- ratings[1:600000,]
```


Explore
```{r}
head(ratings)
```



```{r}
hist(ratings$rating)
```


Convert to realRatingMatrix
```{r}
ratingsMatrix <- sparseMatrix(as.integer(ratings$userId), as.integer(ratings$movieId),  x = ratings$rating)
colnames(ratingsMatrix) <- levels(ratings$movieId)
rownames(ratingsMatrix) <- levels(ratings$userId)
mratings <- as(ratingsMatrix, "realRatingMatrix")
```


Explore
```{r}
mratings
```


```{r}
hist(rowCounts(mratings))
```



```{r}
#table(rowCounts(mratings))
```



```{r}
hist(colCounts(mratings))
```
    


```{r}
#table(colCounts(mratings))
```

   
Select Subset 1 and Subset 2

```{r}
(ratingShort <- mratings[rowCounts(mratings) > 90, colCounts(mratings) > 60])
```




Step-6) Check and Remove Empty Lines
```{r}
ratingShort <- mratings[, colCounts(mratings) > 60]
ratingShort <- ratingShort[rowCounts(ratingShort) > 90, ]
ratingShort
```

```{r}
#table(rowCounts(ratingShort))
```



```{r}
#table(colCounts(ratingShort))
```

```{r}
(ratings_movies <- ratingShort[, colCounts(ratingShort) != 0])
```
 


## Evaluating Recommender Techniques
### Evaluating the model
#### Preparing the data for validation
```{r}
n_fold <- 4
items_to_keep <- 25
rating_threshold <- 3
eval_sets <- evaluationScheme(data = ratings_movies, method = "cross-validation", k = n_fold,
                              given = items_to_keep, goodRating = rating_threshold)
```


## Train and Test sets
#### Train/test split
Split the dataset into test and train sets to build the model.

```{r}
train <- getData(eval_sets, "train")
known <- getData(eval_sets, "known")
unknown <- getData(eval_sets, "unknown")
```


#### Evaluate the item-based collaborative filtering recommender
```{r}
model_to_evaluate <- "IBCF"
model_parameters <- NULL
```

Build the model
```{r}
eval_recommender <- Recommender(data = getData(eval_sets,"train"), method = model_to_evaluate,
                                parameter = model_parameters)
```

Specify the number of items to recommend
```{r}
items_to_recommend <- 10
```

Build the matrix with the predicted ratings
```{r}
eval_prediction <- predict(object = eval_recommender, newdata = getData(eval_sets, "known"),
                           n = items_to_recommend, type = "ratings")
class(eval_prediction)
```

Number of movies recommend to each user
```{r}
qplot(rowCounts(eval_prediction)) + geom_histogram(binwidth = 50) +
  ggtitle("Distribution of movies per users")
```


#### Metrics that mesure the accuracy:
- Root mean square error(RMSE)
- Mean squared error (MSE)
- Mean absollute error(MAE)
#### Metrics by user
```{r}
eval_accuracy <- calcPredictionAccuracy(x = eval_prediction, data = getData(eval_sets, "unknown"),
                                        byUser = TRUE)
head(eval_accuracy)
```

Distribution of RMSE by user
```{r}
qplot(eval_accuracy[,"RMSE"]) + geom_histogram(binwidth = 0.1) +
  ggtitle("Distribution of the RMSE by user")
```


#### Accuracy of the whole model
```{r}
calcPredictionAccuracy(x = eval_prediction, data = getData(eval_sets, "unknown"),
                                        byUser = FALSE)
```

### Evaluation by Recommendation
Comparing the recommendations with the purchases having positive rating
```{r}
results <- evaluate(x = eval_sets, method = model_to_evaluate, n = seq(10,100,10))
class(results)
```

Build the ROC curve
```{r}
plot(results, annotate = TRUE, main = "ROC curve")
```

#### Accuracy metrics Precision and Recall
```{r}
plot(results, "prec/rec", annotate = TRUE, main = "Precision-Recall")
```



## Compare the models

```{r}

```


```{r}
models_to_evaluate <- list(IBCF_cos = list(name = "IBCF", param = list(method = "cosine")),
                           UBCF_cor = list(name = "UBCF", param = list(method = "pearson")),
                           SVD = list(name = "SVD", param = list(k=100)),
                           random = list(name = "RANDOM", param = NULL))
```



Number of the recommendations items to test the models
```{r}
n_recommendations <- c(1, 5, seq(10, 100,10))
```

Run and evaluate the model

```{r}
list_results <- evaluate(x = eval_sets, method = models_to_evaluate, n = n_recommendations)
class(list_results)
```


### Identify the most suitable model
Bild the chart displaying ROC curves 
```{r}
plot(list_results, annotate = 1, legend = "topleft")
title("ROC curves")
```



UBCF with pearson correlation has the better Area Under the Curve(AUC). It is the best-performing technique.


Precision-Recall chart
```{r}
plot(list_results, "prec/rec", annotate = 1, legend = "bottomright")
title("Precision-Recall")
```


### Support for Business or User Experience Goal
Effect of a deployed recommder system on Business value
Base on ads, the goal is to increase the time the user spend with the service and engagement as a proxy for retention
We can craete an hybrid content-based collaborative filtering
```{r}
hybrid_model <- HbridRecommender(models_to_evaluate, c(0.30, 0.30, 0.39, 0.01 ))
hybrid_pred <- predict(hybrid_model, newdata = knowm, type="ratings")
(hybrid_acc = calcPredictionAccuracy(hybrid_pred, unknown))
```



### Comparison of the Accuracy
```{r}

```





